---
title: "Bernard_Asante_Missing_Data_Imputation"
author: "Bernard Asante"
date: "2025-02-01"
output:
  html_document:
    toc: true
    toc_float: true 
    toc_depth: 3
---

# **Loading libraries** 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(visdat)
library(naniar)
library(mice)
library(VIM)
```

# **Reading the dataset**

```{r, warning=FALSE}
path = "T:\\Desktop\\USASK\\Year One\\WINTER\\Data Science\\Data\\can_path_data.csv"

data <- read_csv(path)

#data %>% 
  #glimpse()  # Output is too long. 
```

Loading the entire can_path_student_dataset. This dataset has 440 variables(columns)  and 42287 observations(rows)

## Broad overview of missing data in dataset

```{r}
miss_var_summary(data)
```

It can be observed some variables has more than 99 % missing values, at least from the first 10 rows. These variables are probably missing not at random(MNAR). For efficiency seek, I have decided to retain only variables with less that 20% missing values. 

## Missing data overview

```{r}
data_pct <- data %>%
          select(where(~sum(is.na(.x))/length(.x) < 0.20)) 

missing_table <- miss_var_summary(data_pct)

missing_table 

```

The 'missing_table' displayed above shows the percentage of missing data in the dataset after selecting variables with less that 20% missing data. The dataset with variables containing missing data less that 20% has been named "data_pct". Further processes will be performed on this data_pct instead of the entire dataset(data) 

```{r}
# Selecting 7 columns randomly from the dataset due to high computation power required for the entire dataset
health_data <- data %>% 
  
  select(ID,HS_GEN_HEALTH,NUT_VEG_QTY, NUT_FRUITS_QTY,PA_TOTAL_SHORT,DIS_ASTHMA_EVER,DIS_DIAB_EVER)

write.csv(health_data, "health_data.csv",row.names = FALSE)
```

This step is to further reduced the columns in the dataset to increase efficiency. Eleven(8) columns were selected from the data_pct(containing variables of less than 20% missing data). The output was saved as a new comma separated value(csv) file named "health_data.csv"

# **Sliced dataset[health_data]**

```{r}

health_data <- read_csv("health_data.csv")

health_data %>% 
  glimpse()

```

This is to give an overview of the new dataset (health_data.csv) displaying number of row (41187) and columns(7)


```{r}
health_data  %>% 
  head()  # Displaying the first 6 rows of the dataet 

health_data %>% 
  tail()  # Displaying the first 6 rows of the dataet 
```

```{r}
health_data %>% 
  summary()
```
It can be observed that each variable has a missing data (NA's)

# **Missing Data Exploration**

```{r}
missing_columns <- health_data %>% 
  summarise_all(~sum(is.na(.)))
  

missing_columns

```
Now that I know  the missing values for each columns, I want to visualize them on a simple bar graph to see the order frequency of missing values for each column

## Visualization of missing data per columns 

```{r}
missing_columns_long <- missing_columns %>% 
  pivot_longer(cols = HS_GEN_HEALTH:DIS_DIAB_EVER, names_to = "columns", values_to = "counts")

missing_columns_long

missing_plot <- missing_columns_long %>% 
  ggplot(aes(x = reorder(columns,counts), y = counts))+
  geom_segment(aes(x = reorder(columns,counts), yend = counts, xend = reorder(columns,counts), y =0))+
  geom_point(size = 4, colour = "#a13")+
  coord_flip()+
  labs(title = "Bar Graph of missing data in the dataset by columns", x = "Column_names")
  
missing_plot
```
From the graph, it can be seen that PA-TOTAL_SHORT has the highest missing data around 7000 where as HS_GEN_HEALTH has the least values less that 1000. The next step is to calculate for the percentage of missing data in each column usinng the miss_var_summary

## Percentage of the missing data 

```{r}
missing_data <- miss_var_summary(health_data)

missing_data %>% 
  head(5)
```
I know Physical activity variable has the highest missing values of 16.4% this is evident from the bar graph above.

For other variables, the percentage of missing data is rarely a problem , Nut_veg_qty(6.19%), nut_fruits_qty(5.89%), Dis_asthma_Ever(2.98%) and Hs_Gen_health(1.63%). This is likely to make the data Missing Completely at Random(MCAR). However, it is hard to say for Pa_total_short which has a missing data 0f 16.4%, probably Missing at Random(MAR). These are mere assumptions and need to be analyzed further.

The visdat package was used to make some visualization to better understand the pattern of missing data in the dataset.


# **Visdat** 

## vis_dat()

```{r}
data_col_type <- vis_dat(health_data)

data_col_type
```
The gray color represents missing values in the dataset. It can be seen that ID has no missing values. Few missing values was seen with some variables which corresponds to the percentages above and the bar graph. we can see a lot of missing data in the Pa_total_short column.

## Vis_miss()

```{r}
vis_miss(health_data)
```
This is to display only the missing values. The mising percentage can be seen against the names of each variable on the graph. This is same to the one observed using miss_var_summary from naniar package. 


## vis_cor()

```{r}
health_data_numeric <- health_data %>% 
  select(where(is.numeric)) # Selecting only numeric variables 

vis_cor(health_data_numeric)
```
This heatmap is to display the correlation between the numeric variables in the dataset. There were moderate correlation observed between some of the variables; veg quantity and fruits quantity. similar observation is seen between diabetes ever and asthma ever. From theory, missing data in strongly correlated variables can be imputed using values from the other( except MNAR). But in this analysis, my interest is with Physical activity total which has mild correlation with the nutritional and diseases variables and  mild - moderate correlation with hs_gen_health.
## gg_miss_upset

```{r}
gg_miss_upset(health_data)
```
From the gg-miss_upset plot above, we can observe a connection between missing data points in physical activity and the nutritional(NUT) variables, Physical activity NA is connected to all the variables at the intersection size of 334. 

I want to run further test to assess if the missingness in PA_TOTAL_SHORT is dependent on observed values in the other variables. 

# **Determining the type of missingness**

## Little MCAR test from naniar

```{r}
mcar_test(health_data)

```

The mcar test has the null hypothesis: The missing data is completerly at random(MCAR). Alternative hypothesis states that, missing data is not completely at random that is, it can be MAR or MNAR. 

The test had a pvalue less than 0.001. This is a strong evidence to reject the null hypothesis and conclude that the data is not missing completely at random. 

I want to see if missingness in pa_total  is MAR or MNAR using logistic regression to see how the various variables predicts the "NA" variables in pa_total_short. Logistic regression is used here because the missing data "NA" is not continuous but rather "TRUE" or "FALSE" as can be seen in _missing_pa_total_. 

## Logistic Regression Model

```{r}

missing_Pa_total <- is.na(health_data$PA_TOTAL_SHORT) %>% 
  head()

print(missing_Pa_total)  # This displays the columns values as logical; TRUE and FALSE making binomial logistic regression appropriate

glm(is.na(PA_TOTAL_SHORT) ~ HS_GEN_HEALTH + NUT_FRUITS_QTY + DIS_ASTHMA_EVER + DIS_DIAB_EVER, 
    data = health_data, 
    family = binomial) %>% 
  summary()

```
It can be observed that all the variables has a pvalue < 0.001 indicating they are significant predictors of the missingness in the pa_total_short variable. Again this is an assumption from the model and can not be proven with high certainty

## Geom_miss_plot

```{r}
scatter_plot_miss <- ggplot(health_data, aes(x = NUT_FRUITS_QTY, y = PA_TOTAL_SHORT)) + 
                  geom_miss_point(alpha = 0.2) +
                  labs(x = "Fruits", y = "Physical Activity")
plot(scatter_plot_miss)
```
It can be observed that missing decreases as fruit intake  increases above 8. A lot of missing data was also between 0 to 5000 minutes of exercise, however, missingness reduces above 5000 minutes upwards.



# **Shadow matrix**

```{r}
data_shadow <- health_data %>% 
  bind_shadow()

data_shadow %>% 
  head()
```

The data shadow add and indicator matrix to the dataset to display is a value is missing(NA) or not missing(!NA)

# **Data Imputation**


## Summarizing missing data

```{r}

summary(health_data$PA_TOTAL_SHORT)

health_data %>% summarise(
  n_distinct = n_distinct(PA_TOTAL_SHORT),
  mean_data = mean(PA_TOTAL_SHORT,na.rm = TRUE),
  n_miss = sum(is.na(PA_TOTAL_SHORT)),
  median_data = median(PA_TOTAL_SHORT, na.rm = TRUE)
)
```
This data summary includes unique values (n_distinct = 3421), mean (mean_data = 2574.089), missing values (n_miss = 6763), and median (median_data = 1782).

# **Mean Imputation**

```{r}
data_shadow$pa_not_imputed <- data_shadow$PA_TOTAL_SHORT  # Adding a column called pa_not_imputed to maintain pa_total values after imputation

data_shadow <- impute_mean_at(data_shadow, .vars = vars(PA_TOTAL_SHORT))


data_shadow %>% 
  head()
```
This displays the first six(6) rows of the mean imputed physical activity total. 


## Visualizing Density Plots

```{r}
density_mean_imp <- ggplot(data_shadow) +
                geom_density(aes(pa_not_imputed, colour= "pa_not_imputed")) +
                geom_density(aes(PA_TOTAL_SHORT, colour= "PA_TOTAL_SHORT"))
plot(density_mean_imp)
```
It can be seen that the mean imputed has an increased peak  around 2500 minutes compared to the distribution of the not imputed variable. This shows that  the mean imputation has possibly inflated the mean compared to the sample distribution.

## Descriptive comparison of imputed and non imputed dataset

```{r}
median_pa_not_imputed <- median(data_shadow$pa_not_imputed,na.rm = TRUE)

median_pa_not_imputed


median_pa_imputed <- median(data_shadow$PA_TOTAL_SHORT) # "NA" has been removed so need not to add na.rm
median_pa_imputed


```

It can be observed that the median for non imputed pa short column  is 1782 and increased to 2379 after mean imputation. 


## Boxplot of PA_TOTAL_SHORT before mean_imputation


```{r}
ggplot(data_shadow, aes(x  = pa_not_imputed)) +
         geom_boxplot(outlier.shape = NA) +  # Avoid extreme outliers hiding the boxplot
         labs(title = " Total Physical Activity Before Mean Imputation",
              subtitle = "Median around 1700 ",
              y = "Total Physical Activity")+
         coord_flip()+
         theme_minimal()
```

We can see the median for the non imputed variable is around 1700


## Boxplot of PA_TOTAL_SHORT after mean_imputation

```{r, warning=FALSE}

ggplot(data_shadow, aes(x  = PA_TOTAL_SHORT)) +
  geom_boxplot(oulier.shape = NA) +
  labs(title = " Total Physical Activity after Mean Imputation",
       subtitle = "Median around 2500 ",
       y = "Total Physical Activity") +
  coord_flip()+
  theme_minimal()

```




The median value from the graph can not be determined explicitly with full certainty however, it is probably around  2500. Mean imputation can bias estimate from these observations. 


# **MICE Imputation**

```{r}

data_mice <- mice(health_data, m = 5, maxit = 5) # imputing missing data for each variable in the dataset using MICE with 5  imputed datasets , and 5 control iterations.

data_mice_complete <- complete(data_mice, action = "long")

data_mice_complete %>% 
  head() # Displaying first six(6) rows of the MICE imputed dataset

data_mice_complete %>% 
  tail() # Displaying last six(6) rows of the MICE imputed dataset

data_mice$method
```

I want to join the variables in data_mice_complete with health_data. However, they have the same columns name which may be confusing, so I want to rename the data_mice_complete variable. 

## Renaming of data_mice variables 

```{r}
data_mice_complete <- data_mice_complete %>% 
  rename(HS_GEN_HEALTH_imp = HS_GEN_HEALTH,
         NUT_VEG_QTY_imp = NUT_VEG_QTY,
         NUT_FRUITS_QTY_imp = NUT_FRUITS_QTY,
         PA_TOTAL_SHORT_imp = PA_TOTAL_SHORT,
         DIS_ASTHMA_EVER_imp = DIS_ASTHMA_EVER,
         DIS_DIAB_EVER_imp = DIS_DIAB_EVER)

data_mice_complete %>% 
  head()
```


## Joining the two datasets 

```{r}

health_data_c <- full_join( health_data, data_mice_complete, by = join_by("ID")) 

health_data_c %>% 
  head()

```

The data_mice has been full join to the health_data to a common variable - health_data_c. This is to allow for comparison of imputed and not imputed variables 

## Data Visualization 

### Density plot of PA_TOTAL_SHORT

```{r}
summary(health_data_c$PA_TOTAL_SHORT)
summary(health_data_c$PA_TOTAL_SHORT_imp)

density_imp <- ggplot(health_data_c) +
                geom_density(aes(PA_TOTAL_SHORT, colour= "PA_TOTAL_SHORT")) +
                geom_density(aes(PA_TOTAL_SHORT_imp, colour= "PA_TOTAL_SHORT_imp"))
plot(density_imp)

```

A similar distribution can be seen for Pa-total imputed and pa_total_not imputed. 


###  Density plot of PA_TOTAL_SHORT by .imp

```{r}
density_m_5 <- health_data_c %>% 
  mutate(.imp = as.factor(.imp)) %>%   # Converting .imp to a factor 
  ggplot(aes(x = PA_TOTAL_SHORT_imp, colour = .imp)) +
                geom_density()
plot(density_m_5)
```
A similar distribution was seen for each imputation made using MICE. 


The two main dataset that I am using are the health_data which contains missing data and data_mice_complete  which contains imputed values.Now this datasets have been joined to health_data_c. 

Let me summarize all the datasets before I proceed.

## Descriptive summary of Dataset before and after MICE imputation

###  Descriptive summary of health_data before imputation

```{r}
health_data_miss <- miss_var_summary(health_data)

health_data_miss

```

So the health_data variable contains non impyed variables. 

### Summary of dataset befoe MICE imputation

```{r}

health_data %>% 
  summary()

```

we can see NA's across each variable. This shows missing  data point(s) is not imputed. 

### Boxplot of PA_TOTAL_SHORT before MICE_imputation

```{r}

pa_short_plot <- health_data %>% 
  ggplot(aes(PA_TOTAL_SHORT))+
  geom_boxplot()+
  labs(title = "Boxplot of Total physcial activity before MICE imputation",
       subtitle = "Median can be seen around around 1700")+
  coord_flip()

pa_short_plot
```

We can see the median of Physical activity total is around 1700 minutes in the non imputed dataset.

### Summary of Dataset after MICE Imputation 

```{r}
health_data_imputed <- miss_var_summary(data_mice_complete)
health_data_imputed

data_mice_complete %>% 
  summary()


```

It can be seen that there is no missing data in the health_imputed dataset. This means the data points were correctly imputed using MICE.

### Boxplot of PA_TOTAL_SHORT after MICE_imputation

```{r}

pa_short_plot <- data_mice_complete %>% 
  ggplot(aes(PA_TOTAL_SHORT_imp))+
  geom_boxplot()+
  labs(title = "Boxplot of Total physcial activity after MICE imputation",
       subtitle = "Median can be seen around around 1700")+
  coord_flip()

pa_short_plot
```


We can see the median value for physical activity before and after mice imputation are close ~ 1700. Comparing to mean_imputation( which had a median of ~ 2500). we can say mean_imputation can over or in other situations underestimate a sample parameter. 

```{r}
health_data_joined <- miss_var_summary(health_data_c)
health_data_joined %>% 
  head()
```

# **Using KNN Methods**

## Overview of Dataset without imputed missing values

```{r}
health_data %>% 
  head() # displaying the first six(6) rows in the raw dataset(non imputed)

health_data %>% 
  summarise_all(~sum(is.na(.)))

```


## KNN Imputation using "VIM"
 
```{r}

knn_imputed <- VIM::kNN(health_data, k = 3) # K is number of neighbors.

knn_imputed %>% 
  head()

```


### Summary of Dataset after KNN Imputation

```{r}
knn_imputed %>% 
  summary()

```

This displays the summary of the variables after KNN imputation using K= 3 neighbors.  The various parameter in the summary table has been explained at the end of this script, for now I want ot focus on how it affected the meidan of pa_total compared to the non_imputed pa_total variable

### Boxplot of PA_TOTAL_SHORT before KNN_imputation

```{r}

pa_short_plot <- health_data %>% 
  ggplot(aes(PA_TOTAL_SHORT))+
  geom_boxplot()+
  labs(title = "Boxplot of Total physcial activity before MICE imputation",
       subtitle = "Median can be seen around around 1700")+
  coord_flip()

pa_short_plot

```

This is a boxplot of PA_total on non imputed dataset. 


### Boxplot of PA_TOTAL_SHORT after KNN _imputation

```{r}
ggplot(knn_imputed, aes(PA_TOTAL_SHORT))+
  geom_boxplot()+
  coord_flip()+
  labs(title = "Boxplot of PA_TOTAL_SHORT after KNN Imputation using VIM", 
       subtitle = "Median value around 1800 min")
```


This is the boxplot of PA_total after KNN imputation

# **Regression Model**


## Regression model using non_imputed dataset 

```{r}
pa_fruit_model <- lm (PA_TOTAL_SHORT ~ NUT_FRUITS_QTY, data = health_data )

pa_fruit_model %>% 
  summary()
```


## Regression model using mean imputed dataset 

```{r}
pa_fruit_model <- lm (PA_TOTAL_SHORT ~ NUT_FRUITS_QTY, data = data_shadow )

pa_fruit_model %>% 
  summary()
```
We can see that with a unit increase in NUT_FRUITS_QTY, Physical activity increases by 159.5 minutes which is statistically significant(p_value<0.001)


## Regression model using MICE Imputed dataset

```{r}
set.seed(123)

pa_fruits_model <- lm(PA_TOTAL_SHORT_imp ~ NUT_FRUITS_QTY_imp, data = data_mice_complete) #Linear Regression model  using dataset with imputations
summary(pa_fruits_model)


pa_fruits_miss <- data_mice_complete %>% with(lm(PA_TOTAL_SHORT_imp ~ NUT_FRUITS_QTY_imp)) # Pooled Regression 
summary(pa_fruits_miss)
```
We can see from the pooled estimate that with a unit increase in NUT_FRUITS_QTY, Physical activity increases by 195 minutes which is statistically significant(p_value<0.001). 

## Regression model using KNN Imputed dataset

```{r}

set.seed(123)

pa_asthma_model <- lm(PA_TOTAL_SHORT ~ NUT_FRUITS_QTY, data = knn_imputed) #Linear Regression model  using dataset with KNN imputations
summary(pa_asthma_model)


pa_asthma_miss <- knn_imputed %>% with(lm(PA_TOTAL_SHORT ~ NUT_FRUITS_QTY)) # Pooled Regression 
summary(pa_asthma_miss)
```


# **Summary of findings**

## Table of Median Values accross imputation methods

```{r}
library(knitr)

# Create a data frame for the medians
median_table <- data.frame(
  Imputation_Method = c("Not Imputed", "Mean Imputed", "MICE Imputed", "KNN Imputed"),
  Median_Value = c(1782, 2379, 1782, 1812)
)

median_table
```

We can see the median value is between 1700 and around 1800 for Not imputed, MICE and KNN imputed physical activity. However, the median for the mean imputed physical activity column is 2379, which is an "outlier"

## Regression model of PA_TOTAL_SHORT and NUT_FRUITS_QTY

```{r}

regression_table <- data.frame(
  Imputation_Method = c("Not Imputed", "Mean Imputed", "MICE Imputed", "KNN Imputed"),
  Intercept_B = c(2181, 2249, 2181, 2129),
  Coefficient_NUT_FRUITS_QTY = c(193.79, 159, 195, 185),
  P_Value = c("<2e-16", "<2e-16", "<2e-16", "<2e-16")
)

regression_table

```

A significant relationship was observed between fruit consumption and total minutes of physical activity across all models (p < 0.002). However, the mean-imputed regression model revealed a rather minor increase in physical activity minutes (159) per unit increase in fruit consumption.

This may imply that mean imputation may underestimate the true effect of association between a predictor and an outcome 

It can also be shown that the intercepts and coefficients of the Not Imputed, MICE, and KNN imputed methods are identical. This could be because the percentage of missing data in the dataset is insufficient to skew the estimates. However, a considerable difference may be noticed between imputed and non imputed models if the proportion of data loss is high. 














